Is Partner Modeling Mutual?

### 2014-08-27 mutual-modeling.tex
Import of original Word doc is done

Each of the studies that we present in the next section have been design in such a way that, while tractable on a
robot, mutual modelling plays a key role for the task performance: the tasks
require a ``reasonably high" grounding criterion \groundingcriterion,
and all tasks are about building a solution or a representation together.

@@

>>>>

Each of the studies that we present in the next section have been design in such a way that, while tractable on a
robot, mutual modelling plays a key role for the task performance: the tasks
require a ``reasonably high'' grounding criterion \groundingcriterion,
and all tasks are about building a solution or a representation together.

@@

\section{Synthesis}

Out of 5 studies, 3 revealed a correlation between the model peers built about
each other: when A builds an accurate model of B, B also tends to build an
accurate model of A.  In one study, this was the case only in one condition.
The conclusion is hence that modeling is mutual but only in certain conditions,
summarized in Table~\ref{synthesis_table}.

\begin{table*}[h!t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{p{3cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}}
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
\textit{Task}                 & Game in virtual Space            & Game in physical Space          & Building an argument map  & Building a concept map    & Building a concept map       \\
\textit{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
\textit{Shared Editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
\textit{Group Size}           & 2                                & 3                               & 3                         & 3                         & 3                            \\
\textit{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\textit{Awareness tool}       & Partner's position / orientation & Partner's current/past position & -                         & Partner's concept map     & Partner's scores at pre-test \\
\textit{Dependent variable}   & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge gains & Team/Ind. knowledge gains    \\
\textit{Independent variable} & Awareness tool vs not.           & Awareness tool vs not.          & Audio only vs Audio+Video & Script vs  not            & Awareness tool vs not.       \\
\textit{MM°-DV correlation}   & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
\textit{$\Delta_1$}                   & +0.38 (Sig)                      & +0.41 (Sig)                     & ???                       & +0.26 (near SIG)          & +0.05 (non SIG)              \\
\textit{$\Delta_2$}                   &                                  & +0.30                           & +0.43                     &                           &                              \\
\textit{$\Delta_3$}                   &                                  & ???                             & +0.66                     &                           &                              \\
\textit{$\Delta_4$}                   &                                  &                                 &                           & +0.05                     &                             
\end{tabular}
}
\caption{Comparison of the studies}
\label{synthesis_table}
\end{table*}

Coming back to our initial hypotheses, the $\Delta_2$ results in studies 1, 2
and 3 reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The
\emph{interaction hypothesis} $\mathcal{H}_{3}$ was that the activity of
modeling one's partner is not simply the result of an individual cognitive
effort but overall the side-effect of the quality of interactions. The
\emph{second-order-modeling hypothesis} $\mathcal{H}_{2}$ stated that, if A is
accurate in modeling B, A should notice when B misunderstands A and should be
able to repair. A good modeler would somehow drive the modeling of his or her
partner.  This \emph{drive} will however occur through verbal cues and repairs
and is hence not easy to differentiate from $\mathcal{H}_{3}$.  The
contradictory $\Delta_1$ results in studies 4 (to some extent) and 5 weaken the
hypothesis.

The triangular questions lead to similar contradictions. The $\Delta_2$
correlation between the \gmodel{A}{C} and \gmodel{B}{C} (studies 2 and 3?)
support $\mathcal{H}_{2}$ if modeling skills were purely individual, there would
be no reason for modeling skills of A and B to be correlated.  The role of C in
\gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (studies
3 and 2?) question $\mathcal{H}_{3}$: if the role of the modellee was so
important, it would lead be surprising to have by chance a correlation between
B's and C' ability to drive A's modeling.

In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
($\Delta_2$ in study 4 and study 5) could also contributes to the idea that A's
modeling skills is not the key factor. However, between modeling someone and
modeling oneself, \ie  metacognition, they are so many differences, that we
would hardly conclude from that,

Among these complex results, it is important to stress that, for $\Delta_2$, the
significant correlations are all below .50, not around .90.  So even if all
$\Delta_2$ correlations were significant in all experiments, we would still not
conclude that modelling purely reflects the quality of social interactions.
There is obviously a large part of individual variance; some of us are much
better in modelling peers. This is not new. The contribution of these studies is
rather to establish the importance of team processes on something that a priori
seemed to be more of a personal cognitive skill.  In other words, we stress that
the bottle is half-full.

These conclusions must be presented with multiple disclaimers. First, they
heavily rely on correlations; hence we cannot identify causal links. Second, we
faced hard methodological issues. Providing learners with on-task questionnaires
introduce a bias: they will pay more attention to their partners in the
remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
and rationalization biases. The nature of mutual modeling implies methodological
challenges that call for new  measurement methods. We have promising results for
using eye tracking methods to address this challenge.

Finally, let us stress the expressions we used, such as \gmodel{A}{B}, do not
imply we have a mechanical view of modeling. It's just a language for referring
to different components of modeling each other in teams.  This simple notation
emerges from the need to compare five mutual modelling studies; it did not
generate a consistent program of consistent studies.  Our results are not very
robust because they emerge from a \emph{post-hoc} reinterpretation of studies
that addressed different research questions (media richness, awareness tools,
scripts,...). This diversity makes our results difficult to integrate, partly
contradictory. Nonetheless, this diversity gives some generalizability: mutual
modelling has been investigated in different contexts (virtual space versus real
space), with different groups sizes (pairs and triads) and different tasks.
Moreover, these results have been found using different methods for measuring
modeling: on-task in study 1 versus off-task in study 2, subjective validation
in study 1 versus objective validation in study 2 (comparing A's model with B's
behavior).

\section*{Acknowledgments}

The experiments were developed with the help of Patrick Jermann,  Fabien
Girardin, Marc-Antoine Nüssli, Thomas Werhle, Yvan Bourquin and  Jeremy Goslin.
The main funding has been provided from a NSF Grant grant \#102511-106940.
